---
output:
  md_document:
    variant: markdown_github
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, echo = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "README-"
)
```

# MCHT

*Version `r packageVersion("MCHT")`*

**MCHT** is a package implementing an interface for creating and using Monte
Carlo tests. The primary function of the package is `MCHTest()`, which creates
functions with S3 class `MCHTest` that perform a Monte Carlo test.

## Installation

**MCHT** is not presently available on CRAN. You can download and install 
**MCHT** from GitHub using [**devtools**](https://github.com/r-lib/devtools) via
the R command `devtools::install_github("ntguardian/MCHT")`.

## Monte Carlo Hypothesis Testing

Monte Carlo testing is a form of hypothesis testing where the $p$-values are
computed using the empirical distribution of the test statistic computed from
data simulated under the null hypothesis. These tests are used when the
distribution of the test statistic under the null hypothesis is intractable or
difficult to compute, or as an exact test (that is, a test where the
distribution used to compute $p$-values is appropriate for any sample size, not
just large sample sizes).

Suppose that $s_n$ is the observed value of the test statistic and large values
of $s_n$ are evidence against the null hypothesis; normally, $p$-values would be
computed as $1 - F(s_n)$, where $F(s_n) = F(S_n \leq s_n)$ is the cumulative
distribution functions and $S_n$ is the random variable version of $s_n$. We
cannot use $F$ for some reason; it's intractable, or the $F$ provided is only
appropriate for large sample sizes.

Instead of using $F$ we will use $\hat{F}_N$, which is the empirical CDF of
the same test statistic computed from simulated data following the distribution
prescribed by the null hypothesis of the test. For the sake of simplicity in
this presentation, assume that $S$ is a continuous random variable. Now our
$p$-value is $1 - \hat{F}_N(s_n)$, where $\hat{F}_N(s_n) = \frac{1}{N}\sum_{j =
1}^{N}\mathbb{1}_{\{\tilde{S}_{n,j} \leq s_n\}}$ and $\tilde{S}_{n,j}$ is an
independent random copy of $S_n$ computed from simulated data with a sample size
of $n$.

The power of these tests increase with $N$ (see [1]) but modern computers are
able to simulate large $N$ quickly, so this is rarely an issue. The procedure
above also assumes that there are no nuisance parameters and the distribution of
$S_n$ can effectively be known precisely when the null hypothesis is true (and
all other conditions of the test are met, such as distributional assumptions). A
different procedure needs to be applied when nuisance parameters are not
explicitly stated under the null hypothesis. [2] suggests a procedure using
optimization techniques (recommending simulated annealing specifically) to
adversarially select values for nuisance parameters valid under the null
hypothesis that maximize the $p$-value computed from the simulated data. That is
the procedure employed here. (In fact, the tests created by `MCHTest()` are the
tests described in [2].)

See the documentation for more details and references.

## Examples

`MCHTest()` is the main function of the package and can create functions with S3
class `MCHTest` that perform Monte Carlo hypothesis tests.

For example, the code below creates the Monte Carlo equivalent of a $t$-test.

```{r}
library(MCHT)
library(doParallel)

registerDoParallel(detectCores())  # Necessary for parallelization, and if not
                                   # done the resulting function will complain
                                   # on the first use

ts <- function(x, mu = 0) {sqrt(length(x)) * (mean(x) - mu)/sd(x)}
sg <- function(x, mu = 0) {
  x <- x + mu
  ts(x)
}
rg <- rnorm

mc.t.test <- MCHTest(ts, sg, rg, seed = 20181001, test_params = "mu", 
                     lock_alternative = FALSE,
                     method = "Monte Carlo One Sample t-Test")
```

The object `mc.t.test()` is an S3 class, and a callable function.

```{r}
class(mc.t.test)
```

`print()` will print relevant information about the construction of the test.

```{r}
mc.t.test
```

Once this object is created, we can use it for performing hypothesis tests.

```{r}
dat <- c(2.3, -0.13, 1.42, 1.51, 3.43, -0.96, 0.59, 0.62, 1.28, 4.07)

t.test(dat, mu = 1, alternative = "two.sided")  # For reference
mc.t.test(dat, mu = 1)
mc.t.test(dat, mu = 1, alternative = "two.sided")
```

This is the simplest example; `MCHTest()` can create more involved Monte Carlo
tests. See other documentation for details.

## Planned Future Features

* Permutation tests are a form of Monte Carlo test; perhaps we could add a class
  for constructing permutation tests like we did for Monte Carlo tests.
* Bootstrap tests are yet another form of Monte Carlo tests; we could give some
  constructor for making bootstrap tests too.

## References

1. A. C. A. Hope, *A simplified Monte Carlo test procedure*, JRSSB, vol. 30
   (1968) pp. 582-598
2. J-M Dufour, *Monte Carlo tests with nuisance parameters: A general approach
   to finite-sample inference and nonstandard asymptotics*, Journal of
   Econometrics, vol. 133 no. 2 (2006) pp. 443-477
